---
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Power Calculations for Regression

[![R-CMD-check](https://github.com/zrmacc/Power/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/zrmacc/Power/actions/workflows/R-CMD-check.yaml)

Zachary R. McCaw <br>
Updated: 2026-03-02


### Linear regression

Consider the linear model $Y = X\beta_{X} + Z\beta_{Z} + \epsilon$, with X, Y, and Z centered. The function `PowerLinReg` computes the power to reject $H_{0}:\beta_{X} = 0$ using the usual Wald test.

**Arguments:** `beta_x` (true coefficient for X), `cov_xz` (covariance of X and Z), `n` (sample size), `t1e` (type I error), `var_resid` (residual variance of $Y \mid (X,Z)$), `var_x` and `var_z` (variances of X and Z). When X or Z is multivariate, supply `cov_xz`, `var_x`, and `var_z` as matrices.

```{r}
library(Power)
PowerLinReg(
  beta_x = 1,
  cov_xz = 0.5,
  n = 10,
  t1e = 0.05,
  var_resid = 1,
  var_x = 1,
  var_z = 1
)
```

To find the sample size needed for a target power, use `SampleSizeLinReg` (with `max_n` as the upper bound for the search):

```{r}
SampleSizeLinReg(
  beta_x = 1,
  cov_xz = 0.5,
  max_n = 100,
  power = 0.90,
  t1e = 0.05,
  var_resid = 1,
  var_x = 1,
  var_z = 1
)
```

### Logistic regression

For the logistic model $\mathrm{logit}(\pi) = \beta_0 + X\beta_X + Z\beta_Z$ with $\pi = P(Y=1 \mid X, Z)$, the function `PowerLogisticReg` computes the power to reject $H_0: \beta_X = 0$ using the asymptotic Wald test. The distribution of $(X, Z)$ is assumed multivariate normal with mean zero and covariance given by `var_x`, `var_z`, and `cov_xz`. The expected information is approximated by Gauss–Hermite quadrature (default when dimension is small) or Monte Carlo.

**Arguments:** `beta_0` (intercept), `beta_x` and `beta_z` (true coefficients), `var_x`, `var_z`, `cov_xz` (covariance of X and Z), `n` (sample size), `t1e` (type I error). Use `method = "quadrature"` or `"montecarlo"` to choose the approximation; `n_quad` and `n_sim` control accuracy.

```{r}
PowerLogisticReg(
  beta_0 = 0,
  beta_x = 0.5,
  beta_z = 0.2,
  var_x = 1,
  var_z = 1,
  cov_xz = 0.2,
  n = 200,
  t1e = 0.05,
  method = "quadrature",
  n_quad = 15L
)
```

To find the sample size for a target power, use `SampleSizeLogisticReg`:

```{r}
SampleSizeLogisticReg(
  beta_0 = 0,
  beta_x = 0.5,
  beta_z = 0.2,
  var_x = 1,
  var_z = 1,
  cov_xz = 0.2,
  max_n = 500,
  power = 0.85,
  t1e = 0.05,
  method = "quadrature",
  n_quad = 15L
)
```
