---
title: "Power Calculations"
author: "Zachary McCaw"
date: "2020-12-22"
output: 
  html_document: 
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Linear Regression

Consider the linear model: $Y = X\beta_{X} + Z\beta_{Z} + \epsilon$. Suppose each of X, Y, and Z has been centered to have mean zero. The function `PowerLinReg` determines the power to reject $H_{0}:\beta_{X} = 0$ via the standard Wald test. For example:

```{r}
library(Power)
PowerLinReg(
  beta_x = 1,
  cov_xz = 0.5,
  n = 10,
  t1e = 0.05,
  var_resid = 1,
  var_x = 1,
  var_z = 1
)
```

Here `beta_x` is the true coefficient for X, `cov_xz` is the covariance between X and Z, `n` is the sample size, `t1e` is the type I error, `var_resid` is the residual variance of $Y|(X,Z)$, i.e. the variance of $\epsilon$, `var_x` and `var_y` are the variance of X and Y. Note that when either X or Z is a vector, rather than a scalar, `cov_xz`, `var_x`, and `var_z` should be supplied as matrices.

To determine the necessary sample size for a target power:
```{r}
SampleSizeLinReg(
  beta_x = 1,
  cov_xz = 0.5,
  max_n = 100,
  power = 0.90,
  t1e = 0.05,
  var_resid = 1,
  var_x = 1,
  var_z = 1
)
```

Here `max_n` is an upper bound on the sample size, and `power` is the target power.